The Design and Implementation of the Synchronous Language \CEU

### 2016-09-30 paper.tex
(>) quantitative comparison with Esterel

In previous work~\cite{ceu.sensys13}, we evaluate the implementation of \CEU
in comparison to a code base%
\footnote{TinyOS repository: \url{http://github.com/tinyos/tinyos-release/}}
in \emph{nesC} (a \C variant)~\cite{wsn.nesc}.
%
Figure~\ref{fig.eval} compares source size (\emph{number of tokens}), binary 
size (\emph{ROM}), and memory usage (\emph{RAM}) for a number of standardized 
network protocols and a radio driver.
The small overhead in resource usage shows that the gains in productivity and 
safety with \CEU make it a viable alternative to \C in the context of 
constrained embedded systems.
%
Unfortunately, most real applications in Esterel seem to be closed source.
In future work, we plan to cooperate with groups that have more experience 
and access to real code bases in Esterel for a comprehensive comparison with 
\CEU.
% TODO-no:
%Still, considering \C as an ``absolute benchmark'', we consider the small 
%overhead .  solid results

\begin{figure*}[b]
%\includegraphics[width=\textwidth,clip=true,trim=-80px 385px 300px 40px]{eval}
\caption{ Resource usage for \CEU and \emph{nesC} in the domain of sensor 
networks.
\label{fig.eval}
}
\end{figure*}

@@

Single-threaded dispatching may not be suitable for hard real-time activities.
In previous work~\cite{ceu.sensys13}, we measure how synchronous lengthy 
computations in \C (e.g., hashing and compression) can block the scheduler and 
affect higher-priority activities such as a radio driver.
In such cases, the system requires careful testing to avoid undersized hardware 
deployment.
For instance, we currently do not perform any worst-case reaction times 
analysis~\cite{esterel.wcrt,esterel.fixed}.

@@

Esterel has different compilation back ends that synthesizes to software and 
also to hardware circuits~\cite{esterel.emperor,esterel.tutorial}.
Among the software-based approaches, \emph{SAXO--RT}~\cite{esterel.saxort} is 
the closest to our implementation with respect to trail allocation and 
scheduling:
the compiler slices programs into ``control points'' (analogous to our ``entry 
points'') and rearranges them into a directed acyclic graph respecting the 
constructive semantics of Esterel.
Then, it flattens the graph into sequential code in \C suitable for static 
scheduling.

>>>>

@@

In previous work~\cite{ceu.sensys13}, we evaluate our implementation by
rewriting some components of the \emph{TinyOS} code base%
\footnote{TinyOS repository: \url{http://github.com/tinyos/tinyos-release/}}
from \emph{nesC} (a \C variant)~\cite{wsn.tos,wsn.nesc} to \CEU.
%
TinyOS provides open-source industrial-level applications and serve as basis
for a considerable amount of research in the Wireless Sensor Network academic
community.
%
The generated code has been tested in real hardware and also in congested
networks simulated in sofware~\cite{ceu.sensys13}.
%
Figure~\ref{fig.eval} compares source size (\emph{number of tokens}), binary 
size (\emph{ROM}), and memory usage (\emph{RAM}) for a number of standardized 
network protocols and a radio driver.
The small overhead in resource usage shows that the gains in productivity and 
safety with \CEU make it a viable alternative to \C in the context of 
constrained embedded systems.
%
Single-threaded dispatching may not be suitable for hard real-time activities.
We also measure how synchronous lengthy 
computations in \C (e.g., hashing and compression) can block the scheduler and 
affect higher-priority activities such as a radio driver.
In such cases, the system requires careful testing to avoid undersized hardware 
deployment.
For instance, we currently do not perform any worst-case reaction times 
analysis~\cite{esterel.wcrt,esterel.fixed}.

In future work, we plan to perform an equivalent throughout analysis in
comparison to industrial-quality code bases in Esterel.
%
An Esterel implementation that compiles to sequential C code with similar
techniques also generates ``code less than $10\%$ larger the original
manually-sequenced code'' for an industrial-size
application~\cite{esterel.efficient}.
%
%Unfortunately, most real applications in Esterel seem to be closed source.
%In future work, we plan to cooperate with groups that have more experience 
%and access to real code bases in Esterel for a comprehensive comparison with 
%\CEU.
% TODO-no:
%Still, considering \C as an ``absolute benchmark'', we consider the small 
%overhead .  solid results

\begin{figure*}[b]
\includegraphics[width=\textwidth,clip=true,trim=-80px 385px 300px 40px]{eval}
\caption{ Resource usage for \CEU and \emph{nesC} in the domain of sensor 
networks.
\label{fig.eval}
}
\end{figure*}

@@

Esterel has different compilation back ends that synthesizes to software and 
also to hardware circuits~\cite{esterel.emperor,esterel.tutorial}.
Among the software-based approaches,
\emph{SAXO--RT}~\cite{esterel.saxort,esterel.efficient} is the closest to our
implementation with respect to trail allocation and scheduling:
the compiler slices programs into ``control points'' (analogous to our ``entry 
points'') and rearranges them into a directed acyclic graph respecting the 
constructive semantics of Esterel.
Then, it flattens the graph into sequential code in \C suitable for static 
scheduling.
